---
title: "Initial Korean covid 19"
author: "Nathaly Ruiz"
date: "Fall 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

$H_O$ There is no difference in dying depending on sex          
$A_O$ There is a difference in dying depending on sex
         
result~sex     
Categorical~categorical       
we will use the helper file CAT~CAT   


## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
          
The barchart shows that more males died than females. This shows that dying is dependent on sex. 

  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
          
The numerical results shows that amoung the people who dies there was 64% of males who died and 36% of females who died.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
         
The inferential results shows that the chi-squared test has a p-value of 0.001. And the Fishers exact test has a p-value of 0.00009. Because both of the p-values are under the acceptable rate of 0.05 we will reject the null hypothesis. The Fishers Exact test also states that men are 2.3 more times to die of Covid19 in Korea than women are.


## Conclusions

$H_O$ There is no difference in dying depending on sex         
$A_O$ There is a difference in dying depending on sex
         
This data is from Korea in 2022          
result~sex     
Categorical~categorical       
we will use the helper file CAT~CAT       

We Will reject the null hypothesis because...           
The barchart shows that more males died than females. This shows that dying is dependent on sex.    
The numerical results shows that amoung the people who dies there was 64% of males who died and 36% of females who died.               
The inferential results shows that the chi-squared test has a p-value of 0.001. And the Fishers exact test has a p-value of 0.00009. Because both of the p-values are under the acceptable rate of 0.05 we will reject the null hypothesis. The Fishers Exact test also states that men are 2.3 more times to die of Covid19 in Korea than women are.         